%Parallel programming represents a challenge to computer scientists as the number of possible behaviors is quite large. This number grows exponentially when the program employs a distributed database with a weak level of isolation as Causal Consistency (CC). Therefore, efficient techniques for verifying and testing are required. In this paper we present a complete, sound and optimal algorithm which employs dynamic partial order reduction (DPOR) techniques for exploring all possible histories a transactional concurrent program may have; the only one up to now designed for this purpose. Our algorithm can be used under several isolation levels such as Causal Consistency, Read Atomic or Read Committed with linear memory consumption in all cases. 
%Dynamic Partial Order Reduction for Checking Correctness Against Weak Isolation Levels

Modern applications, such as social networking systems and e-commerce platforms are centered around using large-scale databases for storing and retrieving data. Accesses to the database are typically enclosed in transactions that allow  computations on shared data to be isolated from other concurrent computations and resilient to failures. Modern databases trade off isolation for performance. The weaker the isolation level, the more behaviors a database is allowed to exhibit and it is up to the developer to ensure that their application can tolerate those behaviors. 

In this work, we propose stateless model checking algorithms for studying correctness of such applications that rely on dynamic partial order reduction. These algorithms work for a number of widely-used weak isolation levels, including Read Committed, Causal Consistency, Snapshot Isolation, and Serializability. We show that it is complete, sound and optimal, and runs with polynomial memory consumption in all cases. We report on an implementation of this algorithm in the context of Java Pathfinder applied to a number of challenging applications drawn from the literature of distributed systems and databases.  